Domain-specific working memory loads selectively increase negative interpertations of surprised facial expressions

### 2019-08-19 Manuscript/Manuscript.Rmd
remove analyses
Settled on subj ratings and MD overall for the two MERP findings.

### force normality ###
# RTz.long <- gather(data, key = Condition, value = RTz,
#                     lo.neu.sur_RTz,
#                     hi.neu.sur_RTz,
#                     lo.emo.sur_RTz,
#                     hi.emo.sur_RTz)

# ### run Friedman's test and store results ###
# RTz.friedman <- friedman.test(RTz ~ Condition|subjID, data = RTz.long)
# lo.emo.v.lo.neu.RTz <- wilcox_test(data$lo.emo.sur_RTz, data$lo.neu.sur_RTz, paired = TRUE, conf.int = TRUE)
# lo.emo.v.hi.neu.RTz <- wilcox_test(data$lo.emo.sur_RTz, data$hi.neu.sur_RTz, paired = TRUE, conf.int = TRUE)
# lo.emo.v.hi.emo.RTz <- wilcox_test(data$lo.emo.sur_RTz, data$hi.emo.sur_RTz, paired = TRUE, conf.int = TRUE)
# hi.emo.v.lo.neu.RTz <- wilcox_test(data$hi.emo.sur_RTz, data$lo.neu.sur_RTz, paired = TRUE, conf.int = TRUE)
# hi.emo.v.hi.neu.RTz <- wilcox_test(data$hi.emo.sur_RTz, data$hi.neu.sur_RTz, paired = TRUE, conf.int = TRUE)
# lo.neu.v.hi.neu.RTz <- wilcox_test(data$hi.neu.sur_RTz, data$lo.neu.sur_RTz, paired = TRUE, conf.int = TRUE)
# 
# # RTz.long$RTz <- ave(RTz.long$RTz, RTz.long$subjID, FUN=scale)
# 
# # wide <- spread(RTz.long, key = Condition, value = RTz)
# # hapiro.test(wide$lo.emo.sur_RT)
### make factors ###
# RTz.long$Load <- ifelse(RTz.long$Condition == "lo.neu.sur_RTz", "Low",
#                           ifelse(RTz.long$Condition == "lo.emo.sur_RTz", "Low",
#                                  ifelse(RTz.long$Condition == "hi.neu.sur_RTz", "High",
#                                         ifelse(RTz.long$Condition == "hi.emo.sur_RTz", "High", ""))))
#  
# RTz.long$Type <- ifelse(RTz.long$Condition == "lo.neu.sur_RTz", "Neutral",
#                           ifelse(RTz.long$Condition == "lo.emo.sur_RTz", "Emotional",
#                                  ifelse(RTz.long$Condition == "hi.neu.sur_RTz", "Neutral",
#                                         ifelse(RTz.long$Condition == "hi.emo.sur_RTz", "Emotional", ""))))
#  
# ### make w/in subjs factors ###
# RTz.long$subjID <- as.factor(RTz.long$subjID)
#  
# ### traditional ANOVA ###
# RTaov <- with(RTz.long,
#                     aov(RTz ~ (Load * Type) +
#                         Error(subjID / (Load * Type))), contrasts = contr.sum())
# summary(RTaov)
# RTaov <- tidy(RTaov)

@@

### traditional ANOVA ###
RTaov <- with(data.long,
                   aov(ReactionTime ~ (Load * Type) +
                       Error(subjID / (Load * Type))), contrasts = contr.sum())
summary(RTaov)
RTaov <- tidy(RTaov)
```
To assses for normality, the data were first tested with Shapiro-Wilk's test. All conditions appeared to be sampled from normal distributions (p's > .300). A Load (Low, High) X Type (Neutral, Emotional) repeated measures ANOVA was used to assess differences in the RTs. There was a trend towards a main effect of type, F(`r (RTaov$df[which(RTaov$term == "Type")])`, `r (RTaov$df[which(RTaov$stratum == "subjID")])`) = `r printnum(RTaov$statistic[which(RTaov$term == "Type")])`, p = `r printnum(RTaov$p.value[which(RTaov$term == "Type")])`. There was no effect of domain, F(`r (RTaov$df[which(RTaov$term == "Load")])`, `r (RTaov$df[which(RTaov$stratum == "subjID")])`) = `r printnum(RTaov$statistic[which(RTaov$term == "Load")])`, p = `r printnum(RTaov$p.value[which(RTaov$term == "Load")])`, nor any interaction between load and domain, F(`r (RTaov$df[which(RTaov$term == "Load:Type")])`, `r (RTaov$df[which(RTaov$stratum == "subjID")])`) = `r printnum(RTaov$statistic[which(RTaov$term == "Load:Type")])`, p = `r printnum(RTaov$p.value[which(RTaov$term == "Load:Type")])`.

```{r plot RT data, include = TRUE}
ggplot(data.long, aes(Condition, `ReactionTime`, fill = as.factor(Condition))) +
  geom_bar(stat = "summary", fun.y = "mean") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = .1)
```

``` {r RT by ratings, include = FALSE}
### assess for normality ###
shapiro.test(data$lo.neu.sur_n_RT) ## non-normal
shapiro.test(data$lo.neu.sur_p_RT) ## normal
shapiro.test(data$hi.neu.sur_n_RT) ## non-normal
shapiro.test(data$hi.neu.sur_p_RT) ## normal
shapiro.test(data$lo.emo.sur_n_RT) ## non-normal
shapiro.test(data$lo.emo.sur_p_RT) ## non-normal
shapiro.test(data$hi.emo.sur_n_RT) ## barely normal
shapiro.test(data$hi.emo.sur_p_RT) ## non-normal

### t-test for pos ratings vs. neg ratings
lo.neu.RT.pvn <- wilcox.test(data$lo.neu.sur_n_RT, data$lo.neu.sur_p_RT, paired = TRUE)
wilcox.test(data$hi.neu.sur_n_RT, data$hi.neu.sur_p_RT, paired = TRUE)
wilcox.test(data$lo.emo.sur_n_RT, data$lo.emo.sur_p_RT, paired = TRUE)
wilcox.test(data$hi.emo.sur_n_RT, data$hi.emo.sur_p_RT, paired = TRUE)

### make long ###
lo.neu.resp <- gather(data, key = "Condition", value =  "ReactionTime",
                    lo.neu.sur_n_RT, lo.neu.sur_p_RT)
lo.emo.resp <- gather(data, key = "Condition", value =  "ReactionTime",
                    lo.emo.sur_n_RT, lo.emo.sur_p_RT)
hi.neu.resp <- gather(data, key = "Condition", value =  "ReactionTime",
                    hi.neu.sur_n_RT, hi.neu.sur_p_RT)
hi.emo.resp <- gather(data, key = "Condition", value =  "ReactionTime",
                    hi.emo.sur_n_RT, hi.emo.sur_p_RT)

lo.neu.resp$Condition <- as.factor(lo.neu.resp$Condition)
lo.emo.resp$Condition <- as.factor(lo.emo.resp$Condition)
hi.neu.resp$Condition <- as.factor(hi.neu.resp$Condition)
hi.emo.resp$Condition <- as.factor(hi.emo.resp$Condition)

lo.neu.resp.plot <- ggplot(lo.neu.resp, aes(Condition, ReactionTime, fill = as.factor(Condition))) +
  geom_bar(stat = "summary", fun.y = "mean", na.rm = TRUE) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = .1)

lo.emo.resp.plot <- ggplot(lo.emo.resp, aes(Condition, ReactionTime, fill = as.factor(Condition))) +
  geom_bar(stat = "summary", fun.y = "mean", na.rm = TRUE) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = .1)

hi.neu.resp.plot <- ggplot(hi.neu.resp, aes(Condition, ReactionTime, fill = as.factor(Condition))) +
  geom_bar(stat = "summary", fun.y = "mean", na.rm = TRUE) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = .1)

hi.emo.resp.plot <- ggplot(hi.emo.resp, aes(Condition, ReactionTime, fill = as.factor(Condition))) +
  geom_bar(stat = "summary", fun.y = "mean", na.rm = TRUE) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = .1)

# lo.neu.resp.plot

# hi.neu.resp.plot

#lo.emo.resp.plot

# hi.emo.resp.plot

```
Next, we tested for reaction time differences between positive and negative interpretations of surprised facial expressions within each working memory load condition. A large proportion of the conditions were not normally distributed, thus we used paired-sample Wilcoxon signed rank tests for assessing differences between positive and negative ratings of surprise RTs. This difference was significant *only* for the low emotional load trials, with surprise rated as positive taking signficantly longer than surprise rated as negative (`r printp(lo.neu.RT.pvn$p.value)`). All other comparisons were not significant (p's > .200). Further, this effect survived Bonferroni correction (p = `r printp(.05 / 4)`).

@@

``` {r MAD resp, include = FALSE}
### assess for normality ###
shapiro.test(data$lo.neu.sur_n_MAD) ## normal
shapiro.test(data$lo.neu.sur_p_MAD) ## non-normal
shapiro.test(data$hi.neu.sur_n_MAD) ## barely normal (.058)
shapiro.test(data$hi.neu.sur_p_MAD) ## normal
shapiro.test(data$lo.emo.sur_n_MAD) ## non-normal
shapiro.test(data$lo.emo.sur_p_MAD) ## barely normal (.068)
shapiro.test(data$hi.emo.sur_n_MAD) ## non-normal
shapiro.test(data$hi.emo.sur_p_MAD) ## normal

### t-test for pos ratings vs. neg ratings
wilcox.test(data$lo.neu.sur_n_MAD, data$lo.neu.sur_p_MAD, paired = TRUE) ### different 
wilcox.test(data$hi.neu.sur_n_MAD, data$hi.neu.sur_p_MAD, paired = TRUE) ### trending
wilcox.test(data$lo.emo.sur_n_MAD, data$lo.emo.sur_p_MAD, paired = TRUE) ### non-sig
wilcox.test(data$hi.emo.sur_n_MAD, data$hi.emo.sur_p_MAD, paired = TRUE) ### non-sig

t.test(data$lo.neu.sur_n_MAD, data$lo.neu.sur_p_MAD, paired = TRUE) ### different 
t.test(data$hi.neu.sur_n_MAD, data$hi.neu.sur_p_MAD, paired = TRUE) ### trending
t.test(data$lo.emo.sur_n_MAD, data$lo.emo.sur_p_MAD, paired = TRUE) ### non-sig
t.test(data$hi.emo.sur_n_MAD, data$hi.emo.sur_p_MAD, paired = TRUE) ### non-sig

### make long ###
lo.neu.resp.MAD <- gather(data, key = "Condition", value =  "MAD",
                    lo.neu.sur_n_MAD, lo.neu.sur_p_MAD)
lo.emo.resp.MAD  <- gather(data, key = "Condition", value =  "MAD",
                    lo.emo.sur_n_MAD, lo.emo.sur_p_MAD)
hi.neu.resp.MAD  <- gather(data, key = "Condition", value =  "MAD",
                    hi.neu.sur_n_MAD, hi.neu.sur_p_MAD)
hi.emo.resp.MAD  <- gather(data, key = "Condition", value =  "MAD",
                    hi.emo.sur_n_MAD, hi.emo.sur_p_MAD)

lo.neu.resp.MAD$Condition <- as.factor(lo.neu.resp.MAD$Condition)
lo.emo.resp.MAD$Condition <- as.factor(lo.emo.resp.MAD$Condition)
hi.neu.resp.MAD$Condition <- as.factor(hi.neu.resp.MAD$Condition)
hi.emo.resp.MAD$Condition <- as.factor(hi.emo.resp.MAD$Condition)

lo.neu.resp.MAD.plot <- ggplot(lo.neu.resp.MAD, aes(Condition, MAD, fill = as.factor(Condition))) +
  geom_bar(stat = "summary", fun.y = "mean", na.rm = TRUE) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = .1)

lo.emo.resp.MAD.plot <- ggplot(lo.emo.resp.MAD, aes(Condition, MAD, fill = as.factor(Condition))) +
  geom_bar(stat = "summary", fun.y = "mean", na.rm = TRUE) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = .1)

hi.neu.resp.MAD.plot <- ggplot(hi.neu.resp.MAD, aes(Condition, MAD, fill = as.factor(Condition))) +
  geom_bar(stat = "summary", fun.y = "mean", na.rm = TRUE) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = .1)

hi.emo.resp.MAD.plot <- ggplot(hi.emo.resp.MAD, aes(Condition, MAD, fill = as.factor(Condition))) +
  geom_bar(stat = "summary", fun.y = "mean", na.rm = TRUE) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = .1)

lo.neu.resp.MAD.plot
lo.emo.resp.MAD.plot
hi.neu.resp.MAD.plot
hi.emo.resp.MAD.plot
```
# Discussion
The effect of high vs. low load is still not apparent in these data, just like Mattek et al. 2016. An alternative explanation is that the high load manipulation is not sufficiently difficult to recruit the targeted cognitive resources; however, future work will be needed to better test this alternative.

>>>>

@@

### traditional ANOVA ###
RTaov <- with(data.long,
                   aov(ReactionTime ~ (Load * Type) +
                       Error(subjID / (Load * Type))), contrasts = contr.sum())
summary(RTaov)
RTaov <- tidy(RTaov)
```
To assses for normality, the data were first tested with Shapiro-Wilk's test. All conditions appeared to be sampled from normal distributions (p's > .300). A Load (Low, High) X Type (Neutral, Emotional) repeated measures ANOVA was used to assess differences in the RTs. There was a trend towards a main effect of domain, F(`r (RTaov$df[which(RTaov$term == "Type")])`, `r (RTaov$df[which(RTaov$stratum == "subjID")])`) = `r printnum(RTaov$statistic[which(RTaov$term == "Type")])`, p = `r printnum(RTaov$p.value[which(RTaov$term == "Type")])`. There was no effect of load, F(`r (RTaov$df[which(RTaov$term == "Load")])`, `r (RTaov$df[which(RTaov$stratum == "subjID")])`) = `r printnum(RTaov$statistic[which(RTaov$term == "Load")])`, p = `r printnum(RTaov$p.value[which(RTaov$term == "Load")])`, nor any interaction between load and domain, F(`r (RTaov$df[which(RTaov$term == "Load:Type")])`, `r (RTaov$df[which(RTaov$stratum == "subjID")])`) = `r printnum(RTaov$statistic[which(RTaov$term == "Load:Type")])`, p = `r printnum(RTaov$p.value[which(RTaov$term == "Load:Type")])`.

```{r plot RT data, include = TRUE}
ggplot(data.long, aes(Condition, `ReactionTime`, fill = as.factor(Condition))) +
  geom_bar(stat = "summary", fun.y = "mean") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = .1)
```
Next, we tested for reaction time differences between positive and negative interpretations of surprised facial expressions within each working memory load condition. A large proportion of the conditions were not normally distributed, thus we used paired-sample Wilcoxon signed rank tests for assessing differences between positive and negative ratings of surprise RTs. This difference was significant *only* for the low emotional load trials, with surprise rated as positive taking signficantly longer than surprise rated as negative (`r printp(lo.neu.RT.pvn$p.value)`). All other comparisons were not significant (p's > .200). Further, this effect survived Bonferroni correction (p = `r printp(.05 / 4)`).

@@

# Discussion
The effect of high vs. low load is still not apparent in these data, just like Mattek et al. 2016. An alternative explanation is that the high load manipulation is not sufficiently difficult to recruit the targeted cognitive resources; however, future work will be needed to better test this alternative.
